---
title: "SANBI Species Lists"
author: "Fiona Spooner"
date: "3 September 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
library(here)
library(dismo)
library(dplyr)
library(knitr)
library(reshape2)
library(raster)
library(rgbif)
library(rgdal)
library(sf)
library(tidyr)
library(vroom)
```

The input here is a combined list of all the species listed in the groups of interest in the SANBI Atlases https://www.sanbi.org/resources/atlases/

```{r}
sanbi <- do.call(rbind,lapply(list.files(path = here::here("SANBI/Species_List/"), full.names = TRUE, pattern = "*.csv"), vroom::vroom))
```

Also inputting all of the species listed with occurences in South Africa, Lesotho and Swazliand 

```{r}

gbif<- vroom::vroom(here::here("GBIF/Species_List/All_GBIF_Species_List.csv"))
```

Just selecting out species that are in the groups we're interested in and then adding a column with this network group in 
```{r}
gs<-gbif %>% 
  filter(!is.na(species) & numberOfOccurrences >=10) %>% 
  dplyr::select(kingdom,phylum, class, order, family, species,numberOfOccurrences) %>% 
  filter(class == "Amphibia" | class == "Arachnida" |class == "Aves" | class == "Cycadopsida" | class == "Dacrymycetes" | class == "Gastropoda" | class == "Insecta" | class == "Lecanoromycetes" |class == "Liliopsida" | class == "Lycopodiopsida" | class == "Magnoliopsida" | class == "Mammalia" | class == "Pinopsida" | class == "Polypodiopsida" | class == "Reptilia") %>% 
  arrange(-numberOfOccurrences) 


gs$network_group<-NA
gs$network_group[gs$class == "Mammalia"]<-"Mammals"
gs$network_group[gs$order == "Lepidoptera"]<- "Lepidoptera"
gs$network_group[gs$order == "Odonata"]<- "Odonata"
gs$network_group[gs$class == "Gastropoda"]<- "Gastropods"
gs$network_group[gs$order == "Hemiptera"]<- "Hemiptera"
gs$network_group[gs$family == "Geotrupidae" | gs$family == "Scarabaeidae"]<- "Dungbeetles"
gs$network_group[gs$order == "Anura"]<- "Frogs"
gs$network_group[gs$order == "Neuroptera"]<- "Lacewings"
gs$network_group[gs$class == "Reptilia"]<- "Reptiles"
gs$network_group[gs$order == "Scorpiones"]<- "Scorpions"
gs$network_group[gs$order == "Araneae"]<- "Spiders"
gs$network_group[gs$species == "Orycteropus afer"]<- "Aardvark"
gs$network_group[gs$class == "Aves"]<- "Birds"
gs$network_group[gs$family == "Viperidae" | gs$family == "Lamprophiidae" | gs$family == "Colubridae" | 
                   gs$family == "Elapidae" | gs$family == "Pythonidae" | gs$family == "Typhlopidae"]<- "Snakes"
gs$network_group[gs$order == "Rodentia"]<- "Rats"


gs$QDSs<-NA

gs<-gs %>% 
  dplyr::select("kingdom", "phylum","class", "order", "family","species", "QDSs", "numberOfOccurrences", "network_group") 

colnames(gs)<-colnames(sanbi)

```

```{r}

all_sp<-rbind(gs, sanbi)

all_sp$Binomial<-as.character(all_sp$Binomial)

all_sp<-all_sp %>% 
  group_by(Binomial) %>% 
  mutate(Records_sum = sum(Records_sum))%>%
  filter(!is.na(network_group) & Order != "Cetacea") %>%
  dplyr::select(Kingdom,Phylum,Class,Order,Family,Binomial, Records_sum, network_group) %>% 
  distinct() %>% 
  arrange(-Records_sum)%>% 
  ungroup() 

```

Getting the bird diet data from the eltontraits db (Wilman, 2014) and including this in the networkm group column - doing it a bit clumsily but it's not v important right now.

```{r}

bird_traits<-read.csv(here::here("Traits/birddiet.csv"))

bird_diet<-bird_traits[bird_traits$Scientific %in% all_sp$Binomial,]

bird_diet<-bird_diet[,c("Diet.5Cat","Scientific")]

all_sp<-merge(all_sp, bird_diet, by.y = "Scientific", by.x = "Binomial", all = TRUE)

all_sp$Diet.5Cat[is.na(all_sp$Diet.5Cat)]<-""

all_sp$network_group<-paste(all_sp$network_group, all_sp$Diet.5Cat)

```

For the mammal data I picked out the predominant diet groups and included this in the network_group variable. There was an already column existing for this in the bird data set. The mammal data is also from the eltontraits database. 

```{r}

mamm_traits<-read.csv(here::here("Traits/mammdiet_elton.csv"))

mamm_diet<-mamm_traits[mamm_traits$Scientific %in% all_sp$Binomial,]

md<-mamm_diet[,c("Scientific", "Diet.Inv", "Diet.Vend", "Diet.Vect", 
             "Diet.Vfish", "Diet.Vunk","Diet.Scav", "Diet.Fruit",
             "Diet.Nect", "Diet.Seed", "Diet.PlantO")]

melt_diet<-melt(md)

main_diet<-melt_diet %>% 
  group_by(Scientific) %>%
  filter(value == max(value)) %>% 
  summarise(diet_all = paste(variable, collapse = ","))


all_sp<-merge(all_sp, main_diet, by.y = "Scientific", by.x = "Binomial", all = TRUE)

all_sp$diet_all[is.na(all_sp$diet_all)]<-""

all_sp$network_group<-paste(all_sp$network_group, all_sp$diet_all)


table(all_sp$network_group)

#write.csv(all_sp, "Network_species_record_counts.csv", row.names = FALSE)

```


Combining with GBIF occurrence records - rather than just the species names we have above

```{r, eval = FALSE}

# gbif <- do.call(rbind,lapply(list.files(path = here::here("GBIF/Occurences/"), full.names = TRUE, pattern = "*.csv"), vroom::vroom))
#write.csv(gbif, "")


gbif_za<-vroom::vroom("D:/Fiona/SHEFS/GBIF/Occurences/0003173-190918142434337_ZA_all.csv", delim = "\t")

gbif_es<-vroom::vroom(here::here("GBIF/Occurences/0004701-190813142620410_ES.csv"), delim = "\t")

gbif_ls<-vroom::vroom(here::here("GBIF/Occurences/0003206-190918142434337_LS.csv"), delim = "\t")


gbif_all<-rbind(gbif_za, gbif_es, gbif_ls)


```


```{r}

gbif_fil <- gbif_all %>%
  dplyr::select(
    kingdom,
    phylum,
    class,
    order,
    family,
    genus,
    species,
    verbatimScientificName,
    decimalLongitude,
    decimalLatitude
  ) %>%
  filter(
    !is.na(species) &
      !is.na(verbatimScientificName) &
      !is.na(decimalLatitude) & !is.na(decimalLongitude)
  ) %>%
  distinct()

# gbif_sp<-filter(gbif_all, !is.na(species) & !is.na(decimalLatitude) & !is.na(decimalLongitude))
# 
# gbif_nsp<-filter(gbif_all, is.na(species) & !is.na(verbatimScientificName) & !is.na(decimalLatitude) & !is.na(decimalLongitude)) #come back to these later if needed 
# 
#mutate(binomial = tolower(verbatimScientificName), binomial = paste(toupper(substr(binomial, 1, 1)), substr(binomial, 2, nchar(binomial)), sep="")) %>% 
#filter(!is.na(decimalLatitude) & !is.na(decimalLongitude)) %>% 
#distinct()

#write.csv(gbif_all, here::here("GBIF/Occurences/ZA_LS_ES_all.csv"))

write.csv(gbif_fil, here::here("GBIF/Occurences/ZA_LS_ES_touse.csv"))

```


```{r}

gbif_fil<-vroom(here::here("GBIF/Occurences/ZA_LS_ES_touse.csv"), delim = ",")

bm<-st_read(here::here("WWF_Ecoregions/Species_Selection_Area_True.shp"))
#bm<-readOGR(dsn = here("WWF_Ecoregions/"), layer = "Species_Selection_Area_True")

e<-st_bbox(bm)
```

Filtering out occurrence points that are outside the extent of the biome we're interested in
```{r}
# gbif_sel<-gbif_fil %>% 
#   filter(decimalLongitude >= e[1] & decimalLongitude <= e[3] & decimalLatitude >= e[2] & decimalLatitude <= e[4])

gb_sel<-gbif_fil %>% 
  #dplyr::select(decimalLongitude, decimalLatitude) %>% 
   filter(decimalLongitude >= e[1] & decimalLongitude <= e[3] & decimalLatitude >= e[2] & decimalLatitude <= e[4]) %>% 
  st_as_sf(.,coords = c("decimalLongitude", "decimalLatitude"), crs = crs(bm)) %>% 
  st_intersection(., bm) %>% 
  dplyr::select(kingdom, phylum, class, order, family, genus,species) %>% 
  mutate(binomial = species) %>%
  dplyr::filter(class == "Mammalia" | class == "Gastropoda" | class == "Reptilia" | class == "Aves" | 
           order == "Lepidoptera" | order == "Odonata" | order == "Hemiptera" | order == "Anura" | 
           order == "Neuroptera" | order == "Scorpiones" | order == "Araneae" | order == "Rodentia" |
           family == "Geotrupidae" | family == "Scarabaeidae" | family == "Viperidae" | family == "Lamprophiidae" |
           family == "Colubridae"  | family == "Elapidae" | family == "Pythonidae" | family == "Typhlopidae" ) %>% 
  distinct()
```

```{r}
in_biome<-over(gb_xy, bm , fn = NULL)
in_biome_wh<-which(complete.cases(in_biome))

gbif_in<-gbif_fil[in_biome_wh,]

in_species<-gbif_in %>% 
  dplyr::select(kingdom, phylum, class, order, family, genus,species) %>% 
  mutate(binomial = species) %>%
  dplyr::filter(class == "Mammalia" | class == "Gastropoda" | class == "Reptilia" | class == "Aves" | 
           order == "Lepidoptera" | order == "Odonata" | order == "Hemiptera" | order == "Anura" | 
           order == "Neuroptera" | order == "Scorpiones" | order == "Araneae" | order == "Rodentia" |
           family == "Geotrupidae" | family == "Scarabaeidae" | family == "Viperidae" | family == "Lamprophiidae" |
           family == "Colubridae"  | family == "Elapidae" | family == "Pythonidae" | family == "Typhlopidae" ) %>% 
  distinct()


in_species$genus<-vapply(strsplit(in_species$binomial," "), `[`, 1, FUN.VALUE=character(1))

in_species$species<-vapply(strsplit(in_species$binomial," "), `[`, 2, FUN.VALUE=character(1))


```

```{r}

in_species$network_group<-NA
in_species$network_group[in_species$class == "Mammalia"]<-"Mammals"
in_species$network_group[in_species$order == "Lepidoptera"]<- "Lepidoptera"
in_species$network_group[in_species$order == "Odonata"]<- "Odonata"
in_species$network_group[in_species$class == "Gastropoda"]<- "Gastropods"
in_species$network_group[in_species$order == "Hemiptera"]<- "Hemiptera"
in_species$network_group[in_species$family == "Geotrupidae" | in_species$family == "Scarabaeidae"]<- "Dungbeetles"
in_species$network_group[in_species$order == "Anura"]<- "Frogs"
in_species$network_group[in_species$order == "Neuroptera"]<- "Lacewings"
in_species$network_group[in_species$class == "Reptilia"]<- "Reptiles"
in_species$network_group[in_species$order == "Scorpiones"]<- "Scorpions"
in_species$network_group[in_species$order == "Araneae"]<- "Spiders"
in_species$network_group[in_species$species == "Orycteropus afer"]<- "Aardvark"
in_species$network_group[in_species$class == "Aves"]<- "Birds"
in_species$network_group[in_species$family == "Viperidae" | in_species$family == "Lamprophiidae" | in_species$family == "Colubridae" | 
                   in_species$family == "Elapidae" | in_species$family == "Pythonidae" | in_species$family == "Typhlopidae"]<- "Snakes"
in_species$network_group[in_species$order == "Rodentia"]<- "Rats"


write.csv(in_species, here::here("GBIF/Occurences/species_in_biome_with_network_group.csv"))


```

```{r}

#all_sp$Genus<-vapply(strsplit(all_sp$Binomial," "), `[`, 1, FUN.VALUE=character(1))

#net_df<-merge(all_sp, in_species, by.x = c("Kingdom","Phylum","Class", "Order", "Family","Genus","Binomial"), by.y = c("kingdom", "phylum", "class", "order", "family","genus","binomial"), all = TRUE)

```


```{r}
in_species<-read.csv( here::here("GBIF/Occurences/species_in_biome_with_network_group.csv"), stringsAsFactors = FALSE)

#birds<-filter(in_species, class == "Aves")

gbif_record<-function(genus, species){

    records<-gbif(genus = genus, species = species, download = FALSE, geo = TRUE)
    records<-
    print(paste(genus, species, records, sep = " "))
    return(records)
  
    }

occ_out<-Map(gbif_record, genus =in_species$genus[in_species$class == "Aves"], species = in_species$species[in_species$class == "Aves"])
occs<-do.call(rbind, occ_out)

in_species$gbif_occs<-occs

#birds$gbif_occs<-occs

```

```{r}

dir_creator<-function(group){
  dir.create(here::here("GBIF/Occurences/", group))
  }

lapply(unique(in_species$network_group),dir_creator)
```


```{r}
# occ_download(
#   "scientificName = Cisticola fulvicapilla", 
#   "decimalLatitude < 0",
#   "decimalLongitude > -29",
#   "decimalLongitude < 53",
#   user = "fspooner",
#   pwd = "GBIFPA$$W0RD",
#   email = "spoonerfiona@gmail.com",
#   format = "SIMPLE_CSV"
#   )


gbif_downloader<-function(sp_binomial, network_group){

  occ_out<-occ_search(
  scientificName = sp_binomial, 
  decimalLatitude = '-90,0',
  decimalLongitude = '-29,53',
  limit = 200000, 
  fields = "minimal",
  year = '2010,2019',
  basisOfRecord = 'HUMAN_OBSERVATION'
  )
  
  occ_out$data<-unique(occ_out$data)
  print(paste(network_group, sp_binomial, sep = " "))
  write.csv(occ_out$data, paste0(here::here("GBIF/Occurences", network_group), "/", gsub(" ", "_",sp_binomial), ".csv"), row.names = FALSE)
  
}

fil_sp<-birds %>% 
  filter(gbif_occs <= 200000 & network_group == "Birds")  #Excluding birds for now and we aren't 
                                                      #really interested in sp with less than 10 occurrences

#Need to unique excluding the key column
#Need to do a second pass on the downloaded data
#to check there are more than 10 unique values



Map(gbif_downloader, sp_binomial = fil_sp$binomial[7], network_group = fil_sp$network_group[7])



write.csv(fil_sp,  here::here("GBIF/Occurences/species_in_biome_with_network_group_nobird_tenplus.csv"), row.names = FALSE)


```

#Removing files where there are less than 10 unique rows - excluding the key column

#First pass removing those with less than 10 records

```{r}

ng_dirs<-list.dirs(here::here("GBIF/Occurences/"))[2:15]

all_sp_files<-list.files(ng_dirs, full.names = TRUE)
length(all_sp_files)
```

```{r}
small_file_remove<-function(file){

  rows<-nrow(fread(file, select = 1L))
  
  if (rows < 10){
    file.remove(file)
  }

  print(file)  
}

sf_rm_out<-lapply(all_sp_files, small_file_remove)
```

```{r}
all_sp_files<-list.files(ng_dirs, full.names = TRUE)
length(all_sp_files)
```


#Second pass removing species with less than 10 unique records
```{r}

unique_remove<-function(file){

  csv_open<-read.csv(file, fill = TRUE, header = TRUE)
  
  unique_rows<-as.numeric(
    csv_open %>% 
    dplyr::select(scientificName, decimalLatitude, decimalLongitude) %>% 
    distinct() %>% 
    tally()
  )
  
 if (unique_rows < 10){
    file.remove(file)
  }

  print(file)  
}

sf_rm_out<-lapply(all_sp_files, unique_remove)

all_sp_files<-list.files(ng_dirs, full.names = TRUE)
length(all_sp_files)


```

```{r}
#Third pass looking at sp where the points are well distributed?





```
